### 2018-12-26 ITiCSE2019.tex
More NCSC stuff

%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
\documentclass[sigconf]{acmart}
\usepackage{paralist}
\usepackage{url}
\usepackage[hyphenbreaks]{breakurl}

\def\UrlBreaks{\do\/\do-}

%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.

%%%% Proceedings format for SIGPLAN conferences 
% \documentclass[sigplan, anonymous, review]{acmart}

%%%% Proceedings format for SIGCHI conferences
% \documentclass[sigchi, review]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\copyrightyear{2019}
\acmYear{2019}
\setcopyright{acmlicensed}
\acmConference[ITiCSE '19]{The 24th Annual Conference on Innovation
  and Technology in Computer Science Education}{Jul. 15-17,
  2019}{Aberdeen, UK}
%\acmBooktitle{}
%\acmPrice{15.00}
%\acmDOI{10.1145/3159450.3159547}
%\acmISBN{978-1-4503-5103-4/18/02}
% This slight change to the code may also save 1 or 2 lines of space.

% removes the headers from each page per the preparation instructions, as these are not needed and will be updated with the chairs' actual session names during the pagination/indexing process:
\fancyhead{}

\begin{document}
\title{How Well do we Teach Cybersecurity?\\(Abstract 20 January, Paper 27th)}
%\title{What Do We Mean By Cybersecurity?}  Tom's original was this
%\titlenote{}
%\subtitle{Extended Abstract}
%\subtitlenote{}
\author{James H. Davenport}
\orcid{0000-0002-3982-7545}
\affiliation{%
  \institution{University of Bath}
  \streetaddress{}
  \city{Bath} 
  \country{United Kingdom}
}
\email{j.h.davenport@bath.ac.uk}

\author{Tom Crick}
\orcid{0000-0001-5196-9389}
\affiliation{%
  \institution{Swansea University}
  \streetaddress{}
  \city{Swansea} 
  \country{United Kingdom}
}
\email{thomas.crick@swansea.ac.uk}

\author{Alastair Irons}
\orcid{}
\affiliation{%
  \institution{Sunderland University}
  \streetaddress{}
  \city{Sunderland} 
  \country{United Kingdom}
}
\email{alastair.irons@sunderland.ac.uk}

\author{Tom Prickett}
\orcid{}
\affiliation{%
  \institution{Northumbria University}
  \streetaddress{}
  \city{Newcastle upon Tyne} 
  \country{United Kingdom}
}
\email{tom.prickett@northumbria.ac.uk}

% The default list of authors is too long for headers}
%\renewcommand{\shortauthors}{Moller and Crick}

\begin{abstract}
Abstract here...
\end{abstract}

\keywords{TBC}

\maketitle

\section{Introduction}

Cybersecurity has been in the news for several years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.

There are calls for education to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}.

Nevertheless, it is one thing to write curricula and requirements, and
another thing to deliver appropriate education, and one could
reasonably ask how well this is done in practice.
\subsection{CyberSecurity: for all or for specialists?}
In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.

AI: GDPR essentially expects CyberSecurity to be in all coding. 
JHD: PCI DSS insists on a Web Application Firewall, which is only needed if you don't trust your applications.

Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}. However, given the shortage of specialists, in practice a good generalist with some CyberSecurity expertise can often get a ``specialist'' job.
\subsection{Research Questions}
\begin{enumerate}
\item The person in teh street --- probably not.
\item The general CS graduate.
\item The general CS masters graduate.
\item The specialist CS graduate.
\item The specialist CS masters graduate.
\end{enumerate}
\section{Policy Context}

Substantial CS/digital curriculum reform across the
UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
-- but what about focus on cybersecurity?

UK national economic skills priority
e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
(especially \cite{JCNSS2018a}
and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
\url{https://www.cybersecuritychallenge.org.uk/}).

There is large-scale media
attention on the ``cybersecurity skills gap''
e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
and
\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
etc -- link to UK Digital Strategy and UK Industrial Strategy.

\section{Challenges}

% should we frame this as one or more case studies?
% do we have any national survey data?

\subsection{PCI DSS}
\cite{PCI2018b}.

Alastair to write about their Credit Fraud experience (MSc CyberSecurity).
HE Academy grant (to be referenced).

What should be in the general curriculum.

\subsection{Educational Resources: SQL Injection}
It is 15 years since \cite{Guimaraesetal2004} 
wrote ``All the topics listed above should be presented in the first
Database Course'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}. SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}). Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.

Clearly such a major weakness should be well-taught. One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cyber-security course.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.

Hence \cite{Drop2019} analyzed the database textbooks used by  44 of the top 50 Computer Science
departments in the United States (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). There were seven such boks, but three accounted for the 36 of the 44. Five of the seven (30 of the 44) had no mention of SQL injection. On the other two, the more popular one has a seriously flawed discussion\footnote{``However, they imply that using parameters is equivalent to using a function to add escape characters
around user input. This is incorrect, as using parameters allows
SQL statements to be pre-compiled, and prevents any user input
from being interpreted as code, while escaping user input is not
recommended as a sole defense since imperfect escape functions
can easily be subverted.'' \cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``However, the fact that the first
example should not be used is not discussed until two pages after
the example in the text, and is not mentioned at all in the caption or
on the page where the figure appears. This means a student who is
skimming the text looking for an example to modify for their own
code could simply copy the code that first appears in the example,
without being aware that this is in fact an example of what they
shouldn't do.'' \cite{Drop2019}}.

\subsection{Informal Resources}
The web abounds with informal resources: tutorials and code snippets. How good are these, and how good are people at using these? This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries. Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in WASP's Top Ten \cite{OWASP2017a}.} weaknesses. Searching for these fragments in PHP projects GitHub found 82 instances of these fragments, of which 117 were verified by manual to be vulnerable --- 80\% of which were vulnerable to SQL injection.

\subsection{Staff}
JHD possibly to write about his recruitment experiences.  ? more IoC input?

Alastair's experiences.

? Poll CPHC, and/or survey job adverts.
\subsection{Accreditation}
?AI/TP to write something.

For a few years now there has been a recognized need to build knowledge, skills and capacity in the area of Cybersecurity. This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy. \cite{UKCabinetOffice} or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.

The teaching of Cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of Cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}. There has  been a debate as to whether Cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}. The consensus now is that Cybersecurity is both a discipline in its own right and that Cybersecurity should be taught within Computer Science and related degrees. There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite{ACM2013a} and specialised Cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.

In the United Kingdom, Higher Education provision addresses both approaches. A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and Cybersecurity ( and closely related fields Computer Security, Digital / Computer Forensics, etc). In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to Cybersecurity for entry in September 2019. An even larger number of HEI provide study opportunities related to more general Computer Science. UCAS lists 246 provides for undergraduate programmes related to Computer Science.

Accreditation has evolved to directly addresses the Cybersecurity challenges in both general Computer Science programmes and specialist Cybersecurity programmes. In the UK, accreditation in the broad computing area is being performed by a few different agencies. These include:

\textbf{1. Not for Profit Organisations}
Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements. One of the curricula elements is related to Cybersecurity. Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.  Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes. As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist Cybersecurity education.

The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK. This could potentially include Cybersecurity related skills. Like the Tech Partnership the focus is upon job market requirements.  Given the size of this initiative this potentially has a significant role to play however at time of writing it is not clear what that shall be.

\textbf{2. National Cyber Security Centre (NCSC)}
The NCSC is a UK Government organisation tasked with enhancing the Cybersecurity of the UK. The NCSC publishes and accredits to a number of Cybersecurity standards \cite{NCSC2018a}. These standard are linked to the ACM recommendations for curricula. To date the major focus of NSSC acreditation has been upon Masters degrees specializing in Cybersecurity. More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in Cybersecurity and Computer Science degrees with a significant Cybersecurity focus \cite{NCSC2018b}.

Currently accredited are 15 Cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters Cybersecurity programmes and 1 Cybersecurity Degree Programme with a further 2 provisionally accredited. Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes. This appears to be a positive initiative that will hopefully further develop over time.

\textbf{3. Professional Bodies}
The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of Cybersecurity discipline areas. The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord and Seoul Accord (do we need to reference these). These memorandums support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.   These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}. Both the IET and The BCS have a long history of expecting coverage environmental factors within the programmes they accredit. The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}. Clearly Cybersecurity has been and continues to be part of these expectations.

This ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.  In this context this is realized as an expectation Cybersecurity is an inclusion in all degrees accredited by the BCS. e.g. the expectation for coverage is true for Computer Science as well as Cybersecurity programmes. Two criteria are expected to be covered by all programmes seeking accreditation. These are \cite{BCS2018a}:

2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices

2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system

Additionally programmes seeking Chartered Information Technology Professional Accreditation also have to cover:

3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security

In the context of BCS accreditation, these requirements imply an exit standard that all students on a programme must be able to demonstrate irrespective of the option choices they have made. This means a HEI applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation . A HEI is expected to provide evidence in the form of programme and module specification documentation and example assessment specifications (coursework and examinations). These criteria and the expectation that they are taught and assessed has been present for a number of years.

Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis. Building on the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}. This development was a partnership between the BCS,  The Council of Heads and Professors of Computing, (ISC)2 (the largest not-for-profit membership body of certified information and software security professionals worldwide, with over 130,000 members) and the Office for Cyber Security and Information Assurance in the Cabinet Office. This project lead to the agreement of a set of Cybersecurity Principles and Learning Outcomes for Computer Science and related IT Degrees in the UK \cite{CPHCISC2}. These guidelines are closely tied to  the ACM recommendations for curricula \cite{ACM2013a}.

The agreed Cybersecurity Principles and Learning Outcomes were included in the BCS Guidelines for accreditation in 2016. This resulted in an extended expectation of the coverage of Cybersecurity in programmes accredited by the Institute. The expectation is now that all students upon all accredited programmes will graduate with skills, knowledge and experience related to:
\begin{enumerate}
\item Information and risk
\item Threats and attacks
\item Cybersecurity architecture and operations
\item Secure systems and products
\item Cybersecurity management
\end{enumerate}
This was discussed with the wider education community by a road-show lead by CPHC. (Alastair could you add something here?)

This change to the accreditation guidelines is now in an implementation period. The accreditation process adopted by the BCS is cyclic in nature. Formally the cycle is 5 years in duration. The new expectations have been implemented as follows. To ensure continuous accreditation,  accreditation visits are normally scheduled every 5 years. At the time of the next visit in this accreditation cycle accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.  It is anticipate from 2020(?) the expectation will be all accredited programmes have the new expectations fully embedded.

TO DO - Has this worked? Should we ask the BCS accreditation team a favour to see if they will let us know: 
1. Total number of accredited programmes
2. impact of changes to  BCS Accreditation Cybersecurity requirements - maybe number of HEIs with a 90 day related to it or possibly number with "At Threshold"
3. Would this give us enough evidence of some impact? Clearly it is a work in progress but I think some progress has been made?
4. Any things else?

%\section*{Acknowledgements}

\bibliographystyle{ACM-Reference-Format}
\bibliography{ITiCSE2019}

\end{document}