### 2019-06-09 Response.tex
Response should list all the actionable comments

\font\manual=manfnt
\def\dbend{{\manual\char127}}
\def\eqq{{\buildrel?\over=}}
\def\C{\mathbf{ C}}
\def\N{\mathbf{ N}}
\def\Q{\mathbf{ Q}}
\def\R{\mathbf{ R}}
\def\Z{\mathbf{Z}}
\def\F{\mathbf{F}}
\def\softO{\tilde{\cal O}}
\def\O{{\cal O}}
%\def\N{{\bf N}}
%\def\Q{{\bf Q}}
%\def\R{{\bf R}}
%\def\Z{{\bf Z}}
\def\action#1{\hfil\rlap{\bf#1}\break}
\def\pcite#1{[#1]}
\def\Res{\mathop{\rm Res}\nolimits}
\def\r#1#2{``#1''$\rightarrow$``#2''}
\documentclass{article}
\bibliographystyle{alphaurl}
\usepackage[hyphens]{url}
%\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage[show]{ed}
\usepackage{graphicx}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{notation}{Notation}
\newtheorem{example}{Example}
\newtheorem{problem}{Problem}
\def\decision#1{\par{\bf #1}}
\pagestyle{empty}
\begin{document}
\author{Crick/Davenport/Irons/Prickett}
\title{Response to Referees}
\date{June 2019}
\maketitle

\section{Referee 1}
No comments that require addressing in detail.
\section{Referee 2}
No comments that require addressing in detail.
\section{Referee 3}
\begin{enumerate}
\item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
\item[A]We have done what we can, but note that the other referees disagree.
\item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
\item[A]
\item What do the authors mean by ``There has been a recent international working group but this has yet to report.'' 
\item[A]
\item It would add more clarity to introduce what these terms DevOps vs DevSecOps represent in the paper. Also, please use the acronym GDPR after introducing it during the first usage such as General Data Protection Regulation (GDPR). 
\item[A]
\item There are repeated ``and'' in the sentence ``In particular the expectation of Privacy by design...'' Also, the ``by'' after ``and'' is not required.

Please consider rephrasing to ``While both Privacy by design and privacy by defualt have been expected to be good practices''.

\item[A]
\item Please rephrase the following statement ``How might accreditation regard cybersecurity education, or help with it?'' as the sentence structure doesn't make much sense. 
\item[A]
\item It is not clear what is the message the authors are conveying in the 2nd paragraph? It lists three different web links and a reference. Also, the web links in 2nd and 3rd paragraphs could be moved to the reference and they could be referenced with a number in this paragraph. 
Again, please use the acronym KA after introducing it as Knowledge Areas (KA) at the first instance of use.

\item[A]
\item  What is JHD? Please expand?
\item[A]
\item Please expand OWASP. 
\item[A]
\item 
The references to first author in the 2nd paragraph could be avoided and the sentence rephrased to highlight only the facts and not any personal experience. 
\item[A]
\item 
 What does ``three accounted for the 36 of the 44'' mean? Please clarify. I am assuming the 36 and 44 refer to number of universities. Please add that clarification, if it is so.
\item[A]
\item 
``If there are only 82 instances of these fragments, how can 117 of them be verified?'' Please add clarification to this discussion.
\item[A]
\end{enumerate}
\section{Referee 4}
\begin{enumerate}
\item ``The problem is it is not a research paper in even the broadest interpretation of that sense. The RQs are invented based on topic areas the authors wish to riff on, only RQ3 comes (somewhat) close to a RQ with an associated presentation based on the literature and some analysis. The 1st 2 RQs, as well as the Introduction sections, are selectively argued, stream-of-consciousness almost writing. Entertaining, informative, arguable, debatable - but not research. No methodology is applied, no systematic review, etc. Instead a smattering of anecdotal information and quotes (and even more quotes in the footnotes) intended to convince the reader of a viewpoint that is never quite articulated clearly - until the end when it comes down to a small set of specific complaints about SQL Injection, XSS, and StackOverflow (which is overall a bit of a letdown as I felt as if this bloglike writing was building to a more impactful crescendo, some kind of kick-in-the-pants call to action for the community).''
\item[A]
\end{enumerate}
\section{Programme Committee}
No comments that require addressing in detail.

\end{document}
\item 
\item[A]
\item 
\item[A]
\item 
\item[A]

\end{document}
\bibliography{../../../../jhd}
\section{}
\begin{description}
\item
\end{description}
\begin{itemize}
\item
\end{itemize}
\begin{example}
\end{example}
\begin{definition}
\end{definition}
\begin{theorem}
\end{theorem}
\begin{enumerate}
\end{enumerate}
\begin{description}
\item[Theme 1]
\item[Theme 2]
\item[Theme 3]
\item[Theme 4]
\item[Theme 5]
\end{description}